On the discovery of K-enhanced and Mg-depleted stars throughout the Milky Way

### 2018-02-07 article/ms.tex
minor changes to ms, larger changes to plotting code

\begin{abstract}
Stars with unusual elemental abundances offer clues about rare astrophysical events or nucleosynthetic pathways. Stars with significantly depleted magnesium and enhanced potassium (e.g., $[{\rm K}/{\rm Fe}] > 1$; $[{\rm Mg}/{\rm Fe}] < -0.5$) have to date only been found in the massive globular cluster NGC~2419, and to a lesser extent, NGC~2808. The origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. Here we present 112 field stars, identified from \LamostGiants\ \lamost\ giant stars, that show significantly enhanced [K/Fe] and possibly depleted [Mg/Fe] abundance ratios.
Our sample spans a full range of metallicities ($-1.5 < [{\rm Fe}/{\rm H}] < 0.3$), yet none show abundance ratios of [K/Fe] or [Mg/Fe] that are as extreme as those observed in NGC~2419. 
We conclude that the nucleosynthetic process producing the anomalous abundances in [K/Fe] and [Mg/Fe] probably (but not necessarily) occurs at a wide range of metallicities. This would suggest that pollution scenarios that are limited to early epochs (e.g., Population III supernovae) are an unlikely explanation. This sample is expected to help guide modelling attempts to explain the origin of the Mg-K abundance signature, which we show exists throughout the Milky Way.
\end{abstract}

@@

\section{Method \& Observations}
\label{sec:method}
\subsection{Candidate selection from \lamost\ spectra}
We used a set of \LamostGiants\ giant stars from the second \lamost\ data release \todo{\citep{lamostDR2}}. The spectra were placed at rest-frame on a common wavelength sampling from 3905\,\AA\ to 9000\,\AA\ and pseudo-continuum-normalised as per \citet{ho2017}. \tc\ \citep{ness2016,ho2017} was used to estimate effective temperature ($\teff$), surface gravity ($\logg$), metallicity ([Fe/H]), and mean $\alpha$-element abundance relative to iron ([$\alpha$/Fe]) using 9,952 stars in common between \lamost\ and the \apogee\ \citep{alam2015} survey. Unless otherwise stated, these transferred labels are those referred to throughout this study. For details regarding the preparatory work, model generation, and label transfer between the \apogee\ and \lamost\ surveys, we direct the reader to \citet{ho2017}.

@@

Abundance ratios of [$\alpha$/Fe] in our sample fall within the lower limits for [Mg/Fe] of the Mg-normal population of both NGC 2419 and NGC 2808, although as we are unable to decouple [Mg/Fe] from $[\alpha$/Fe], the exact level of Mg depletion for the candidates is unclear. The [Mg/Fe] abundances derived from Magellan/MIKE's high resolution spectra are consistent with the Mg-normal population in NGC 2419. However, the \lamost\ spectra relative to \tc\ model is consistent with an under-abundance of Mg. The small sample of stars for which we have high resolution spectra, coupled with the conflict with the \lamost spectra, means we are unable to confidently make a statement regarding the level of [Mg/Fe] compared to NGC 2419 or NGC 2808.

@@

The [K/Fe] abundance ratios in our sample also show an anti-correlation with total metallicity, as illustrated in Figure \ref{KvsFe}, which is not present in the bulk of Milky Way stars. It is interesting that the levels of [K/Fe] in the Mg-poor population of NGC 2419 appear consistent with this relation. However we caution that this may be an artefact of our analysis, further high resolution observations are needed to verify this trend.

@@

While this scenario is consistent with the total spread in metallicity, it is not applicable to NGC 2419 or NGC2808. SAGB-binaries cannot account for 30\% of the stars in NGC 2419 showing depleted Mg and enhanced K abundances. \todo{Why not, exactly?} Furthermore, it would only irregularly produce a depletion in [Mg/Fe], unlike the apparent bi-modality seen in abundance ratios in NGC 2419 and NGC 2808.

@@

\todo{One major issue facing this explanation is that it requires a single gas cloud to form many generations of stars almost continuously in order to be consistent with Figure \ref{KvsFe}. This is only known to occur in inside true galaxies, which are massive enough to form multiple star forming regions. Allowing the candidate stars to have originated in a dwarf galaxy which later merged with the Milky Way introduces many issues. The required mass to be polluted by the mystery object increases drastically, implying multiple pollution events. But invoking multiple pollution events means that the rarity of the signal is harder to explain: if it happened multiple times within a single dwarf galaxy, why not in other places? Further, a merger should result in stars with highly elliptical orbits, while the candidates identified appear to be disk stars.}

\todo{More generally, the [K/Fe]-[Fe/H] relation we observed is supportive of the origin of the Mg-K anti-correlation being temporally confined to an early epoch in the evolution of galaxies and globular clusters. However, this is not necessarily the only interpretation. For example, it could be that the K forming process is happening throughout the life of the galaxy, but the amount of K produced decreases with the metallicity of the polluter star, though there is no obvious reason why this would be the case.}

@@

We find extreme abundance ratios of [K/Fe] $\approx 1.5$ in our sample, but none that exceed the levels observed in NGC 2419 (up to [K/Fe] $\approx$ 2). While our analysis suggests the presence of substantial [Mg/Fe] depletion, follow-up observations are necessary to quantify the level of Mg-depletion. The spread in metallicities among our sample suggests that the mechanism responsible for the Mg-K abundance signature is not limited to early epochs (e.g., Population III supernova) or low-metallicity environments. We highlight SAGB-binaries as a plausible polluter candidate for the Mg-K abundance signature. We encourage future high-resolution observations and modelling efforts to understand this peculiar nucleosynthetic pathway.

>>>>

\begin{abstract}
Stars with unusual elemental abundances offer clues about rare astrophysical events or nucleosynthetic pathways. Stars with significantly depleted magnesium and enhanced potassium (e.g., $[{\rm K}/{\rm Fe}] > 1$; $[{\rm Mg}/{\rm Fe}] < -0.5$) have to date only been found in the massive globular cluster NGC~2419, and to a lesser extent, NGC~2808. The origin of this abundance signature remains unknown, as does the reason for its apparent exclusivity to these two globular clusters. Here we present 112 field stars, identified from \LamostGiants\ \lamost\ giant stars \todo{using a data-driven model produced by \tc}, that show significantly enhanced [K/Fe] and possibly depleted [Mg/Fe] abundance ratios.
Our sample spans a full range of metallicities ($-1.5 < [{\rm Fe}/{\rm H}] < 0.3$), yet none show abundance ratios of [K/Fe] or [Mg/Fe] that are as extreme as those observed in NGC~2419. 
We conclude that the nucleosynthetic process producing the anomalous abundances in [K/Fe] and [Mg/Fe] probably (but not necessarily) occurs at a wide range of metallicities. This would suggest that pollution scenarios that are limited to early epochs (e.g., Population III supernovae) are an unlikely explanation. This sample is expected to help guide modelling \todo{and observational} attempts to explain the origin of the Mg-K abundance signature, which we show exists throughout the Milky Way.
\end{abstract}

@@

\section{Method \& Observations}
\label{sec:method}
\subsection{Candidate selection from \lamost\ spectra}
We used a set of \LamostGiants\ giant stars from the second \lamost\ data release \citep{luo2016vizier}. The spectra were placed at rest-frame on a common wavelength sampling from 3905\,\AA\ to 9000\,\AA\ and pseudo-continuum-normalised as per \citet{ho2017}. \tc\ \citep{ness2016,ho2017} was used to estimate effective temperature ($\teff$), surface gravity ($\logg$), metallicity ([Fe/H]), and mean $\alpha$-element abundance relative to iron ([$\alpha$/Fe]) using 9,952 stars in common between \lamost\ and the \apogee\ \citep{alam2015} survey. Unless otherwise stated, these transferred labels are those referred to throughout this study. For details regarding the preparatory work, model generation, and label transfer between the \apogee\ and \lamost\ surveys, we direct the reader to \citet{ho2017}.

@@

Abundance ratios of [$\alpha$/Fe] in our sample fall within the lower limits for [Mg/Fe] of the Mg-normal population of both NGC 2419 and NGC 2808, although as we are unable to decouple [Mg/Fe] from $[\alpha$/Fe], the exact level of Mg depletion for the candidates is unclear. The [Mg/Fe] abundances derived from Magellan/MIKE's high resolution spectra are consistent with the Mg-normal population in NGC 2419. However, the \lamost\ spectra relative to \tc\ model is consistent with an under-abundance of Mg. The small sample of stars for which we have high resolution spectra, coupled with the conflict with the \lamost\ spectra, means we are unable to confidently make a statement regarding the level of [Mg/Fe] compared to NGC 2419 or NGC 2808.

@@

The [K/Fe] abundance ratios in our sample also show an anti-correlation with total metallicity, as illustrated in Figure \ref{KvsFe}, which is not present in the bulk of Milky Way stars. It is interesting that the levels of [K/Fe] in the Mg-poor population of NGC 2419 appear consistent with this relation. However we caution that this may be an artefact of our analysis. Further high resolution observations are needed to verify this trend.

@@

While this scenario is consistent with the total spread in metallicity, it is not applicable to NGC 2419 or NGC2808. It would take a highly irregular IMF for SAGB-binaries to account for 30\% of the stars in NGC 2419 showing depleted Mg and enhanced K abundances.  Furthermore, SAGB-binaries would only irregularly produce a depletion in [Mg/Fe], unlike the consistent bi-modality seen in abundance ratios in NGC 2419 and NGC 2808.

@@

%\todo{One major issue facing this explanation is that it requires a single gas cloud to form many generations of stars almost continuously in order to be consistent with Figure \ref{KvsFe}. This is only known to occur in inside true galaxies, which are massive enough to form multiple star forming regions. Allowing the candidate stars to have originated in a dwarf galaxy which later merged with the Milky Way introduces many issues. The required mass to be polluted by the mystery object increases drastically, implying multiple pollution events. But invoking multiple pollution events means that the rarity of the signal is harder to explain: if it happened multiple times within a single dwarf galaxy, why not in other places? Further, a merger should result in stars with highly elliptical orbits, while the candidates identified appear to be disk stars.}

%\todo{More generally, the [K/Fe]-[Fe/H] relation we observed is supportive of the origin of the Mg-K anti-correlation being temporally confined to an early epoch in the evolution of galaxies and globular clusters. However, this is not necessarily the only interpretation. For example, it could be that the K forming process is happening throughout the life of the galaxy, but the amount of K produced decreases with the metallicity of the polluter star, though there is no obvious reason why this would be the case.}

@@

We find extreme abundance ratios of [K/Fe] $\approx 1.5$ \todo{can we say they were aprox 1.5? only 1 star was that high... surely an 'up to' or an 'as high as' is more accurate} in our sample, but none that exceed the levels observed in NGC 2419 (up to [K/Fe] $\approx$ 2). While our analysis suggests the presence of substantial [Mg/Fe] depletion, follow-up observations are necessary to quantify the level of Mg-depletion. The spread in metallicities among our sample suggests that the mechanism responsible for the Mg-K abundance signature is not limited to early epochs (e.g., Population III supernova) or low-metallicity environments. We highlight SAGB-binaries as a plausible polluter candidate for the Mg-K abundance signature \todo{within the disk}. We encourage future high-resolution observations and modelling efforts to understand this peculiar nucleosynthetic pathway.
